Kinematics of the Palomar 5 stellar stream from RR Lyrae stars

### 2019-10-03 paper/main.tex
add dai reference

The Pal 5 stream---at a heliocentric distance of $d \sim 20~\kpc$---was discovered using early SDSS imaging, which reached well below the main sequence turnoff of the stream over a large area surrounding the cluster \citep{Odenkirchen:2001, Rockosi:2002}.
Since then, radial velocities have been measured for a handful of giant branch stars in the cluster and along the stream \citep{Odenkirchen:2002, Odenkirchen:2009, Ibata:2017}, and deeper imaging data has been used to map the tails at higher signal-to-noise \citep{Bernard:2016, Ibata:2016, Bonaca:2019}.
The detailed but still limited kinematic data for Pal 5 has made it a canonical example of tidal destruction and a useful tool for constraining Galactic structure.
For example, using the sky track and radial velocity information, the stream has been used to measure the enclosed mass of the Milky Way \citep{Kuepper:2015, Bovy:2016}.
Using deep photometry, the density variations along the stream have been used to place limits on the abundance of massive substructures in the Galactic halo \citep{Erkal:2017}.

@@

\begin{figure*}[t]
\begin{center}
\includegraphics[width=\textwidth]{members.pdf}
\caption{Sky positions of RRLs with Pal 5 colored by membership probability.
The background greyscale shows the surface density of CMD-filtered main sequence stars from \citet{Bonaca:2019}.
Note that the membership probability does not incorporate sky position information, but a clear over-density of high-probability RRL members do trace the stream on the sky.
Purple lines outline the stream track window defined in \citet{Bonaca:2019} and used here to exclude stream stars when constructing the background model.
\changes{As with \figurename~\ref{fig:trackmembers}, RRLs within the sky track are shown as star markers, and RRLs with probability $<0.1$ are plotted with a smaller marker.}
The inset panel shows a $1.5^\circ$ by $1.5^\circ$ zoom-in around the Pal 5 cluster, with the Jacobi radius of the cluster shown as a large red circle.
}
\label{fig:members}
\end{center}
\end{figure*}

@@

Using a small number of RRLs then thought to be associated with Pal 5, \citet{Vivas:2006} measured a distance to the cluster of $d = 22.3~\kpc$.
However, this used a now imprecise $M_V$--$\feh$ relation to compute the distance modulus.
Adopting $\feh = -1.35$ \citep{Ishigaki:2016} and using the updated RRL $M_V$--$\feh$ relation from \citet{Muraveva:2018}, we find that this past RRL-based distance to Pal 5 is instead $d_{\rm Pal 5} = 21.2 \pm 0.6~\kpc$.
To summarize: Our inferred distance to the Pal 5 cluster, $d_{\rm Pal 5} = \clderr$, is consistent with (but more precise than) all recent distance measurements, once corrected for dust extinction and updated calibrations of the horizontal branch absolute magnitude.

>>>>

The Pal 5 stream---at a heliocentric distance of $d \sim 20~\kpc$---was discovered using early SDSS imaging, which reached well below the main sequence turnoff of the stream over a large area surrounding the cluster \citep{Odenkirchen:2001, Rockosi:2002}.
Since then, radial velocities have been measured for a handful of giant branch stars in the cluster and along the stream \citep{Odenkirchen:2002, Odenkirchen:2009, Ibata:2017}, and deeper imaging data has been used to map the tails at higher signal-to-noise \citep{Bernard:2016, Ibata:2016, Bonaca:2019}.
The detailed but still limited kinematic data for Pal 5 has made it a canonical example of tidal destruction and a useful tool for constraining Galactic structure.
For example, using the sky track and radial velocity information, the stream has been used to measure the enclosed mass of the Milky Way \citep{Kuepper:2015, Bovy:2016, Dai:2018}.
Using deep photometry, the density variations along the stream have been used to place limits on the abundance of massive substructures in the Galactic halo \citep{Erkal:2017}.

@@

\begin{figure*}[t]
\begin{center}
\includegraphics[width=\textwidth]{members.pdf}
\caption{Sky positions of RRLs with Pal 5 colored by membership probability.
The background grayscale shows the surface density of CMD-filtered main sequence stars from \citet{Bonaca:2019}.
Note that the membership probability does not incorporate sky position information, but a clear over-density of high-probability RRL members do trace the stream on the sky.
Purple lines outline the stream track window defined in \citet{Bonaca:2019} and used here to exclude stream stars when constructing the background model.
\changes{As with \figurename~\ref{fig:trackmembers}, RRLs within the sky track are shown as star markers, and RRLs with probability $<0.1$ are plotted with a smaller marker.}
The inset panel shows a $1.5^\circ$ by $1.5^\circ$ zoom-in around the Pal 5 cluster, with the Jacobi radius of the cluster shown as a large red circle.
}
\label{fig:members}
\end{center}
\end{figure*}

@@

Using a small number of RRLs then thought to be associated with Pal 5, \citet{Vivas:2006} measured a distance to the cluster of $d = 22.3~\kpc$.
However, this used a now imprecise $M_V$--$\feh$ relation to compute the distance modulus.
Adopting $\feh = -1.35$ \citep{Ishigaki:2016} and using the updated RRL $M_V$--$\feh$ relation from \citet{Muraveva:2018}, we find that this past RRL-based distance to Pal 5 is instead $d_{\rm Pal 5} = 21.2 \pm 0.6~\kpc$.
\changes{Interestingly, \citet{Dai:2018} also independently infer a nearer distance to the cluster around $d_{\rm Pal 5} \approx 21~\kpc$ from dynamical modeling of the stream.}
To summarize: Our inferred distance to the Pal 5 cluster, $d_{\rm Pal 5} = \clderr$, is consistent with (but more precise than) all recent distance measurements, once corrected for dust extinction and updated calibrations of the horizontal branch absolute magnitude.
